---
title: "Question-4"
output: html_notebook
---

## Classification using Naive Bayes 
Example: Filtering spam SMS messages 
Step 2: Exploring and preparing the data 
Read the sms data into the sms data frame
```{r}
sms_raw <- read.csv("sms_spam.csv", stringsAsFactors = FALSE)
```
Examine the structure of the sms data
```{r}
str(sms_raw)
```
Convert spam/ham to factor.
```{r}
sms_raw$type <- factor(sms_raw$type)
```
Examine the type variable more carefully
```{r}
str(sms_raw$type)
table(sms_raw$type)
```
Build a corpus using the text mining ```(tm)``` package
```{r}
library(tm)
sms_corpus <- VCorpus(VectorSource(sms_raw$text))
```
Examine the SMS Corpus
```{r}
print(sms_corpus)
inspect(sms_corpus[1:2])
as.character(sms_corpus[[1]])
lapply(sms_corpus[1:2], as.character)
```

Clean up the corpus using ```tm_map()```
```{r}
sms_corpus_clean <- tm_map(sms_corpus, content_transformer(tolower))
```
Show the difference between sms_corpus and corpus_clean
```{r}
as.character(sms_corpus[[1]])
as.character(sms_corpus_clean[[1]])
```
Remove numbers
```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, removeNumbers)
```
Remove stop words
```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, removeWords, stopwords())
```
Remove punctuation
```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, removePunctuation)
```
Tip: create a custom function to replace (rather than remove) punctuation
```{r}
removePunctuation("hello...world")
replacePunctuation <- function(x) { gsub("[[:punct:]]+", " ", x) }
replacePunctuation("hello...world")
```
Illustration of word stemming
```{r}
library(SnowballC)
wordStem(c("learn", "learned", "learning", "learns"))
sms_corpus_clean <- tm_map(sms_corpus_clean, stemDocument)
```
Eliminate unneeded whitespace
```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, stripWhitespace)
```
Examine the final clean corpus
```{r}
lapply(sms_corpus[1:3], as.character)
lapply(sms_corpus_clean[1:3], as.character)
```
Create a document-term sparse matrix
```{r}
sms_dtm <- DocumentTermMatrix(sms_corpus_clean)
```
Alternative solution: create a document-term sparse matrix directly from the SMS corpus
```{r}
sms_dtm2 <- DocumentTermMatrix(sms_corpus, control = list(
  tolower = TRUE,
  removeNumbers = TRUE,
  stopwords = TRUE,
  removePunctuation = TRUE,
  stemming = TRUE
))
```
Alternative solution: using custom stop words function ensures identical result
```{r}
sms_dtm3 <- DocumentTermMatrix(sms_corpus, control = list(
  tolower = TRUE,
  removeNumbers = TRUE,
  stopwords = function(x) { removeWords(x, stopwords()) },
  removePunctuation = TRUE,
  stemming = TRUE
))

```
Compare the result
```{r}
sms_dtm
sms_dtm2
sms_dtm3
```
Creating training and test datasets
```{r}
sms_dtm_train <- sms_dtm[1:4169, ]
sms_dtm_test  <- sms_dtm[4170:5559, ]
```

Also save the labels
```{r}
sms_train_labels <- sms_raw[1:4169, ]$type
sms_test_labels  <- sms_raw[4170:5559, ]$type
```
Check that the proportion of spam is similar
```{r}
prop.table(table(sms_train_labels))
prop.table(table(sms_test_labels))
```

Word cloud visualization
```{r}
library(wordcloud)
wordcloud(sms_corpus_clean, min.freq = 50, random.order = FALSE)
```
Subset the training data into spam and ham groups
```{r}
spam <- subset(sms_raw, type == "spam")
ham  <- subset(sms_raw, type == "ham")
wordcloud(spam$text, max.words = 40, scale = c(3, 0.5))
wordcloud(ham$text, max.words = 40, scale = c(3, 0.5))
sms_dtm_freq_train <- removeSparseTerms(sms_dtm_train, 0.999)
sms_dtm_freq_train

```
Indicator features for frequent words
```{r}
findFreqTerms(sms_dtm_train, 5)
```
Save frequently-appearing terms to a character vector
```{r}
sms_freq_words <- findFreqTerms(sms_dtm_train, 5)
str(sms_freq_words)
```
Create DTMs with only the frequent terms
```{r}
sms_dtm_freq_train <- sms_dtm_train[ , sms_freq_words]
sms_dtm_freq_test <- sms_dtm_test[ , sms_freq_words]
```
Convert counts to a factor
```{r}
convert_counts <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

```
Apply ```convert_counts()``` to columns of train/test data
```{r}
sms_train <- apply(sms_dtm_freq_train, MARGIN = 2, convert_counts)
sms_test  <- apply(sms_dtm_freq_test, MARGIN = 2, convert_counts)
```

Step 3: Training a model on the data
```{r}
library(e1071)
sms_classifier <- naiveBayes(sms_train, sms_train_labels)
```
Step 4: Evaluating model performance 
```{r}
sms_test_pred <- predict(sms_classifier, sms_test)
library(gmodels)
CrossTable(sms_test_pred, sms_test_labels,
           prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,
           dnn = c('predicted', 'actual'))
```
Step 5: Improving model performance
```{r}
sms_classifier2 <- naiveBayes(sms_train, sms_train_labels, laplace = 1)
sms_test_pred2 <- predict(sms_classifier2, sms_test)
CrossTable(sms_test_pred2, sms_test_labels,
           prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,
           dnn = c('predicted', 'actual'))
```

