---
title: "Quesion 10"
output: html_notebook
---
## Evaluating Model Performance 
Create the predicted probabilities from the SMS classifier built in Chapter 4.
Obtain the predicted probabilities
```{r}
sms_test_prob <- predict(sms_classifier, sms_test, type = "raw")
head(sms_test_prob)
```
Combine the results into a data frame
```{r}
sms_results <- data.frame(actual_type = sms_test_labels,
                          predict_type = sms_test_pred,
                          prob_spam = round(sms_test_prob[ , 2], 5),
                          prob_ham = round(sms_test_prob[ , 1], 5))
write.csv(sms_results, "sms_results.csv", row.names = FALSE)
```
Confusion matrixes in R
```{r}
sms_results <- read.csv("sms_results.csv")
```

The first several test cases
```{r}
head(sms_results)
```
Test cases where the model is less confident
```{r}
head(subset(sms_results, prob_spam > 0.40 & prob_spam < 0.60))
```
Test cases where the model was wrong
```{r}
head(subset(sms_results, actual_type != predict_type))
```
Specifying vectors
```{r}
table(sms_results$actual_type, sms_results$predict_type)
```

Alternative solution using the formula interface 
```{r}
xtabs(~ actual_type + predict_type, sms_results)
```
Using the CrossTable function
```{r}
library(gmodels)
CrossTable(sms_results$actual_type, sms_results$predict_type)
```
Accuracy and error rate calculation
Accuracy
```{r}
(152 + 1203) / (152 + 1203 + 4 + 31)
```
Error rate
```{r}
(4 + 31) / (152 + 1203 + 4 + 31)
```
Error rate = 1 - accuracy
```{r}
1 - 0.9748201
```
Beyond accuracy: other performance measures
```{r}
library(caret)
confusionMatrix(sms_results$predict_type, sms_results$actual_type, positive = "spam")
```
Kappa statistic example using SMS classifier
```{r}
pr_a <- 0.865 + 0.109
pr_a
pr_e <- 0.868 * 0.888 + 0.132 * 0.112
pr_e
k <- (pr_a - pr_e) / (1 - pr_e)
k
```
Calculate kappa via the vcd package
```{r}
library(vcd)
Kappa(table(sms_results$actual_type, sms_results$predict_type))
```
Calculate kappa via the irr package
```{r}
library(irr)
kappa2(sms_results[1:2])
```
```{r}
sens <- 152 / (152 + 31)
sens
spec <- 1203 / (1203 + 4)
spec
```

Sensitivity and specificity example using SMS classifier example using the caret package
```{r}
library(caret)
sensitivity(sms_results$predict_type, sms_results$actual_type, positive = "spam")
specificity(sms_results$predict_type, sms_results$actual_type, negative = "ham")
```
Precision and recall
```{r}
prec <- 152 / (152 + 4)
prec
rec <- 152 / (152 + 31)
rec
```
Example using the caret package
```{r}
library(caret)
posPredValue(sms_results$predict_type, sms_results$actual_type, positive = "spam")
sensitivity(sms_results$predict_type, sms_results$actual_type, positive = "spam")
```
F-measure
```{r}
f <- (2 * prec * rec) / (prec + rec)
f
f <- (2 * 152) / (2 * 152 + 4 + 31)
f
```
Visualizing Performance Tradeoffs
```{r}
library(pROC)
sms_roc <- roc(sms_results$actual_type, sms_results$prob_spam)
```
ROC curve for Naive Bayes
```{r}
plot(sms_roc, main = "ROC curve for SMS spam filter", col = "blue", lwd = 2, legacy.axes = TRUE)
```
Compare to kNN 
```{r}
sms_results_knn <- read.csv("sms_results_knn.csv")
sms_roc_knn <- roc(sms_results$actual_type, sms_results_knn$p_spam)
plot.new()
plot(sms_roc_knn, col = "red", lwd = 2, add = TRUE)
```
Calculate AUC for Naive Bayes and kNN
```{r}
auc(sms_roc)
auc(sms_roc_knn)
```
Estimating Future Performance and partitioning data
```{r}
library(caret)
credit <- read.csv("credit.csv")
```
Holdout method using random IDs
```{r}
random_ids <- order(runif(1000))
credit_train <- credit[random_ids[1:500],]
credit_validate <- credit[random_ids[501:750], ]
credit_test <- credit[random_ids[751:1000], ]
```
Using caret function
```{r}
in_train <- createDataPartition(credit$default, p = 0.75, list = FALSE)
credit_train <- credit[in_train, ]
credit_test <- credit[-in_train, ]
```
10-fold CV
```{r}
folds <- createFolds(credit$default, k = 10)
str(folds)
credit01_test <- credit[folds$Fold01, ]
credit01_train <- credit[-folds$Fold01, ]
```
Automating 10-fold CV for a C5.0 Decision Tree using ```lapply()```
```{r}
library(caret)
library(C50)
library(irr)
credit <- read.csv("credit.csv")
```
Use an older random number generator to match the book
```{r}
RNGversion("3.5.2") 
set.seed(123)
folds <- createFolds(credit$default, k = 10)

cv_results <- lapply(folds, function(x) {
  credit_train <- credit[-x, ]
  credit_test <- credit[x, ]
  credit_model <- C5.0(default ~ ., data = credit_train)
  credit_pred <- predict(credit_model, credit_test)
  credit_actual <- credit_test$default
  kappa <- kappa2(data.frame(credit_actual, credit_pred))$value
  return(kappa)
})

str(cv_results)
mean(unlist(cv_results))
```

